---
title: "20191008_p8105_casestudy_datawrang2"
author: "Kevin S.W."
date: "10/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

# Practicing a Data Scientist mindset

```{r}

library(tidyverse)
library(dplyr)
library(p8105.datasets)
library(leaflet)

data(nyc_airbnb)

```

This dataset of nyc_airbnb is started by someone who is interested in seeing how airbnb contributes to gentrification.

Some questions to consider:
* Does rating correlate with location?
* How is price related to location?
* How airbnb prices relate to rent in neighborhood
* Which neighborhood is most expensive, etc
* do hosts with multiple sites have higher price/ratings?
* Does price have any relation to ratings?
* average length of stay related to neighborhood? price? etc?

```{r}

test_data <- nyc_airbnb %>% 
  drop_na(review_scores_location) %>% 
  select(review_scores_location, neighbourhood_group, neighbourhood, review_scores_location, price, room_type) %>%
  mutate(
    room_type = as.factor(room_type)
  ) %>% 
  group_by(review_scores_location) %>% 
  ggplot(aes(x = review_scores_location, y = price, color = room_type)) +
  geom_point(alpha = 0.4, aes(size = price)) +
  facet_grid(room_type~neighbourhood_group)

```


# considering spatial locations

```{r}

#rentals located
nyc_airbnb %>%
 ggplot(aes(x = lat, y = long, color = neighbourhood_group)) +
 geom_point()

```

# Using leaflets

```{r}

pal <- colorNumeric(
  palette = "viridis",
  domain = nyc_airbnb$review_scores_location
)

airbnb_map <- nyc_airbnb %>%
  filter(neighbourhood_group == "Manhattan") %>% 
  mutate(
    label = str_c("<b>Stars: ", neighbourhood_group, "</b><br>Price: ", price)
  ) %>% 
  sample_n(2000) %>% # instead of trying to pull x > y, just sample x amount from the data
  leaflet() %>%     # here's the map
  addTiles() %>%    # adding tiles -> here's my map
  #addMarkers(~ lat, ~ long) %>%   # adding markers based on lat/long
  addProviderTiles(providers$CartoDB.Positron) %>% # changing the base map (tile)
  addCircleMarkers(~lat, ~long, 
                   radius = 1, 
                   color = pal(nyc_airbnb$review_scores_location), 
                   popup = ~ label) # modifying the markers

```
